---
title: "Assignment 3 - CRD"
author: "Pramod Hegde"  
format:
  html:
    code-fold: false
    embed-resources: true
    toc: true
    number-sections: false
    theme: cerulean
---

# Introduction

This assignment will ask you to conduct a full CRD analysis, from data import through publication ready plot.

The different sections representing what we did in class were left in the script to provide you a scaffold to start from.

# Instructions

-   Move this script into the `code` subfolder of your `03_crd` project folder\
-   Move the file `wheat_nk_bamyan.csv` into the `data` folder of your `03_crd` project folder\
-   On this script, fill your name under the `title` section of the YAML

# a) Setup

Here is where we load the packages we will use.\
Make sure to load each package in the chunk below.

```{r setup}
#| warning: false
#| message: false
library(tidyverse) 
library(car) 
library(broom) 
library(emmeans) 
library(multcomp) 
library(multcompView)

```

Now, import the data set `wheat_nk_bamyan.csv` in the chunk below.

```{r data import}
df <- read.csv("../Data/wheat_nk_bamyan.csv")


```

The **objective of this study** was to assess the effect of different N and K rates on wheat yield.

For that, a study was conducted using a completely randomized design as the experimental design with four replicates. The treatment design was a 3 N x 3 K rate factorial, for a total of 9 treatments.

# b) EDA tables

Use this space to conduct exploratory data analysis of the imported data, in the form of tables.

```{r}
summary(df)
```

```{r}
glimpse(df)
```

# c) Wrangling

Here, make sure to wrangle your data set with all needed steps. Refer back to class code if you need a refresher on what was needed. As seen on the above table, all data are in dbl format. For better analysis, we need to convert rep, nrate and krate to factor, which will be done by the following chunk.

```{r}
wrangled <- df %>% 
  mutate(rep = factor(rep),
         nrate_kgha = factor(nrate_kgha),
         krate_kgha = factor(krate_kgha),
         trtname = paste0(nrate_kgha, "+", krate_kgha), .before = yield_kgha
         )

glimpse(wrangled)
```

# d) EDA plots

Now, explore the wrangled data using plots.\
Make sure to explore plots showing the main effects and the interaction.

```{r}
wrangled %>% 
  ggplot(aes(nrate_kgha, yield_kgha))+
  geom_boxplot()+
  geom_jitter(aes(color = nrate_kgha))
```

Nothing too discernible in the figure, except for the fact that the yield seems to be increasing with increasing N rates.

```{r}
wrangled %>% 
  ggplot(aes(krate_kgha,yield_kgha))+
  geom_boxplot()+
  geom_jitter(aes(color = krate_kgha))
```

Nothing too discernible in the figure, except for the fact that treatments with 20 KgK/ha have a wider dispersion from the mean, as seen by the height of the box, and the length of the wicks.

```{r}
wrangled %>% 
  ggplot(aes(nrate_kgha,yield_kgha,
             color = nrate_kgha))+
  geom_boxplot()+
  geom_jitter()+
  facet_grid(.~krate_kgha)
```

The first two grids i.e treatments with 0,30 kg of K, have a linear increase in their mean yield with increasing N rates, but the third grid with K treatment of 60 kg/ha, there is a strong jump in yield values with N treatment of 100 kg/ha almost in par with the 200 N kg/ha treatment, implying some interaction effects. 

```{r}
wrangled %>% 
  ggplot(aes(krate_kgha,yield_kgha,
             color = krate_kgha))+
  geom_boxplot()+
  geom_jitter()+
  facet_grid(.~nrate_kgha)
```

There seems to be no interaction effects in N rates of 0 and 100, but we can see some effects when N rate is 200. The yield values for the treatment with 200kg of N and 30kg of K has a higher yield than similar K treatements but with different N rates. The first two grids follow an increasing pattern as K rate increases, which is absent in the third grid signifying interaction effects. 

## Question 1) Below each plot of this section, write a sentence explaining your visual interpretation of the plot.

# e) Statistical model

In this section, run the appropriate model.\
Remember to set the contrast type to sum-to-zero (the default in R is set-to-zero, which is not what we want).

Also, remember to specify all model terms needed for the treatment design.

```{r}
options(contrasts = c("contr.sum", "contr.poly"))

mod <- lm(yield_kgha ~ nrate_kgha + krate_kgha + nrate_kgha:krate_kgha,
              data = wrangled
              )
summary(mod)


```

# f) ANOVA table

Use this section to check the ANOVA table, ensuring you use type 3 sum-of-squares.

```{r}
Anova(mod, type = 3)

```

## Question 2) Based on the ANOVA table above, for which effect(s) in the ANOVA should you extract means and perform pair-wise comparisons? Explain why.

Based on the ANOVA table, there is no significant interaction effects and hence we proceed to testing the effects of individual treatments which are both significant. 

# g) Linear model assumptions

Use this section to first compute your residuals, including studentdized residuals.

## Question 3) What are the linear model assumptions that we need to check for? Create subsections (using ##) below for each of these assumptions, and write code to produce plots to check for these assumptions. For each assumption, write down your interpretation of the produced plots, and your thoughts if the assumptions hold or not.

## Extracting residuals

```{r}
resid <- augment(mod) %>% 
  mutate(.studresid = rstudent(mod))
resid


```

The columns ".std.resid" and ".studresid" are the studentized residuals. The two are the same but calculated by different packages (slightly different methods).

## Model assumption 1 - Residual independence

```{r}
resid %>% 
  ggplot(aes(.fitted,.studresid))+
  geom_point()+
  geom_smooth()


```

There does not seem to be any definite pattern in the residuals, satisfying the first model assumption of residual independennce.

## Model assumption 2 - Residual homoscedascity

Homoscedascity is when we observe an increase in the value of residuals as we move along a gradient. A fan shaped smooth curve indicates heteroscedascity, which is not what we want to see. This assumption can be tested using a similar graph as above.

```{r}
resid %>% 
  ggplot(aes(.fitted, .studresid))+
  geom_point()+
  geom_smooth()


```

We see no fan shaped curve, therefore satisfying the model assumption of homescedascity.

## Model assumption 3 - Residual normality

We expect to see a normal distribution of the residuals centered around 0. Quantile - quantile plot will help us check this assumption

```{r}
resid %>% 
  ggplot(aes(sample = .studresid))+
  stat_qq()+
  stat_qq_line()



```
```{r}
resid %>% 
  ggplot(aes(.studresid))+
  geom_density()+
  scale_x_continuous(limits = c(-3,3), breaks = c(-3:3))

```
From the above two graphs, we can say that the residuals are normally distributed around mean 0. 

## Model assumption 4 - Residual outliers 
Residuals as far as possible should stay within the limits of +- 3 standard deviations from the mean. A fitted vs residuals curve will help us interpret for any outliers. 
```{r}
resid %>% 
  ggplot(aes(.fitted , .studresid))+
  geom_point()+
  geom_smooth()+
  geom_hline(yintercept = c(-3,3), color = "red")


```
None of the residuals are beyond the threshold of 3 SDs, and hence this model assumption is proved. Now that all assumptions have been proved, we can move ahead to using the model for inference. 

# h) Model means

Use this section to extract model means according to your decision of the ANOVA table effects.


From the anova table, we have seen that interaction effects have not been significant, but we will extract means for this. 
```{r}
means_all <- emmeans(mod, 
                ~nrate_kgha * krate_kgha)
means_all
```
This was not really needed because we did not find any interaction in the anova table. 


The following chunk calculates the mean only for N rate. 
```{r}
means_n <- emmeans(mod, ~ nrate_kgha)
means_n

```
```{r}
means_k <- emmeans(mod, ~krate_kgha)
means_k


```
# i) Pairwise comparisons

Use this section to perform pair-wise comparison of the means extracted above.

Make sure to reverse the order from high to low, use letters instead of numbers, and no adjustment.

```{r}
cld_n <- cld(means_n,
              reversed = T,
              Letters = letters,
              adjust = "none")
cld_n
```
```{r}
cld_k <- cld(means_k,
              reversed = T,
              Letters = letters,
              adjust = "none")
cld_k
```
```{r - conversion to a df}
cld_n_converted <- cld_n %>% 
  as.data.frame() %>% 
  mutate(letter = trimws(.group)) 
cld_n_converted
  

```
```{r - conversion to a df2}
cld_k_converted <- cld_k %>% 
  as.data.frame() %>% 
  mutate(letter = trimws(.group))
cld_k_converted
```



# g) Final plot(s)

Produce final plot(s) to show your results.\
Make sure to include both **raw data** and **statistical model summary (i.e., letter separation)**.

Be mindful of your plot to ensure you are creating it in an intuitive way for your audience in relation to interpreting the letter separation structure you used.

In class, we did not have time to make it publication-ready (e.g., a plot looking professional and ready to be included in a manuscript).

Here, make sure you deliver final plot(s) that is/are publication ready.

Make sure to provide an interpretation for the plot(s) created, including which treatment levels were significantly higher/lower.

```{r}
ggplot(mapping = aes(fill = nrate_kgha)) +
  geom_boxplot(data = wrangled,
               aes(nrate_kgha, yield_kgha, fill = nrate_kgha),
               alpha = 0.6
               ) +
  geom_jitter(data = wrangled,
              aes(nrate_kgha, yield_kgha),
              shape = 21,
              alpha = 0.6,
              size = 3)+
  geom_label(data = cld_n_converted,
             aes(nrate_kgha, emmean, label = letter),
             fill = "white"
             )+
  labs(x = "N rate (kg/ha)",
       y = "Yield (kg/ha)")+
  theme_classic()+
  theme(legend.position = "none")+
  scale_fill_viridis_d()

```
From the above graph, we can say that N rate of 0 kg/ha had significantly lower yield. Although the treatments of 100 and 200 kg/ha have slightly different yield values(visually), they are not statistically different. 


```{r}
ggplot(mapping = aes(fill = krate_kgha))+
  geom_boxplot(data = wrangled, 
               aes(krate_kgha, yield_kgha, fill = krate_kgha),
               alpha = 0.6
               )+
  geom_jitter(data = wrangled,
              aes(krate_kgha, yield_kgha),
              shape = 21,
              alpha = 0.6,
              size = 3)+
  geom_label(data = cld_k_converted,
             aes(krate_kgha, emmean, label = letter),
             fill = "white"
             )+
  labs(x = "K rate (kg/ha)",
       y = "Yield (kg/ha)")+
  scale_fill_viridis_d()+
  theme_classic()+
  theme(legend.position = "none")



```
K rate of 0 kg/ha has a significantly lower yield as compared to the other two treatments. 



# Submitting your work

Once you have developed all the code and answers, make sure to Render this quarto file.

**Notes on rendering**:

-   Make sure to render your work and inspect how the final html look like.\
-   If it does not look professional for whatever reason, then fix the issue, re-render it, recheck.\
-   Only send me your work once your html file looks professional.
    -   **DO NOT** delete the file's heading levels (# and ##). They set up the proper heading 1 and 2 levels, and I use them to guide my grading.\

    -   If a given chunk is also outputting warnings or messages, inhibit this behavior by changing the chunk options `message` and `warning` to `FALSE` (I'll demo this in a moment).

    -   If, after rendered, 2 lines of text are connected and you wish to "break line" between them, add 2 extra spaces after the first one (I'lld demo this in a moment).

After rendering, an .html file will be created on your `code` folder.

Rename this file to `Assignment-03-crd-LASTNAME.html`.

For ex., mine would be `Assignment-03-crd-Bastos.html`.

Send the **html** file to **lmbastos\@uga.edu** by Feb 15th 11:59 pm.
